---
title: "Assignment 1"
author: "Lewis White"
date: "2023-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)

source(here("photovoltaic_energy.R"))
```

In this course, we will work a lot with writing and applying functions; You’ve all been exposed to writing functions in previous courses - but just to make sure that everyone remembers the basic syntax in R, complete the following:

Write a function that computes energy produced from a photovoltaic system if you know the average annual solar radiation. You can use the following formula:

E=A∗r∗H∗PR

E is energy (kWh),
A is the solar panel area (m2)
r is panel yield (0-1) (manufacture efficiency - usually around 0.2),
PR is performance ratio (0-1) (accounting for site factors that impact efficiency usually around 0.75)
H is annual average solar radiation (kWh)
Write your function in R; make sure that you add documentation; include default values for r and PR in the function definition; save the function in a single file to upload to Canva.

```{r}
# Function to compute energy produced from a photovoltaic system

# Inputs
# area: solar panel area (m2)
# solar_radiation: annual average solar radiation (kWh)
# yield: panel yield (0-1) (default value: 0.2)
# performance_ratio: performance ratio (0-1) (default value: 0.75)

# Output
# energy: energy produced (kWh)

photovoltaic_energy <- function(area, yield, solar_radiation = 0.2, performance_ratio = 0.75) {
  
  if (area < 0) {
    stop("It looks like you entered a negative value for area. Please try again and make sure your value for area is input correctly.")
  }
  
  if (yield < 0 | yield > 1) {
    stop("It looks like you entered a value for yield that is outside the accepted range. Pnael yield values are between 0 and 1. Please try again and make sure your value for yeild is input correctly.")
  }
  
  if (solar_radiation < 0) {
    stop("It looks like you entered a negative value for area. Please try again and make sure your value for area is input correctly.")
  }
  
  if (performance_ratio < 0 | performance_ratio > 1) {
    stop("It looks like you entered a value for performance ratio that is outside the accepted range (0 - 1). Please try again and make sure your value for the performance ratio is input correctly.")
  }
  
  energy = area * yield * solar_radiation * performance_ratio
  return(energy)
}
```
